Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values

### 2020-08-14 Dissonance_Santiago_v2/Response-to-Reviewers-v2.Rmd

---
title: "Response to Reviewers"
output: pdf_document
---

We would like to thank the editor and three anonymous reviewers for their thoughtful comments and suggestions to revise the paper. We believe that the paper has been substantially improved by their input. In this letter we respond to the feedback received, and describe the actions taken to revise the paper.

\newpage

# Reviewer 1

Policy and aims: The main and most important thing to work on is answering the ‘so what?’ question. Your findings are interesting but how does it change the way we see the world and think about transport? Although, I understand that exploring policy aims in detail is not the focus of your study, because implications for policy and interventions are an important aim of the journal, try and sell the importance of your findings and analysis for practical policies and practice with as much persuasiveness as you can. Are there specific areas of policy that your analysis inform. What policies or practice could be employed in which the differences between age groups, income or commute length, for example, would be necessary to take into account.

>\textcolor{blue}{Added: In various contexts, not only in Santiago, the perception of users is considered mainly from the scope of the trip satisfaction. Understanding dissonances allow public policy to recognise that public transport users are clearly at a disadvantage compared to other transport modes. On the other hand, the results recognise that within public transport users, there are also socioeconomic disadvantages – compared to other groups. This is relevant information for the generation of public policies that promote the use of sustainable modes and public transport, considering aspects of their personal well-being that transcend the areas of travel satisfaction. The factors defined in this study can, and should, be considered in the evaluation of measures that promote the use of sustainable transport in the long term.}

Justifying the use of SWB for transport: I'm not clear about how users can be attracted (line (L)26) to public transport and active travel. would overcoming psychological taboos be enough? So general point, need to argue well for the importance of SWB around travel, in comparison to considerations of travel time, and cost - I know this is hard to do!

>\textcolor{blue}{The consideration of factors related to the subjective personal well-being will enable targeting public policy strategies focused on nudging public and active transport users}

Contribution of the paper to the literature: I don't think you need to number the ways the paper contributes to literature the way you have - you can just say 'Firstly,...'etc. I would actually leave most of the material under these 1,2,3 points untill later in the paper. I would include the bits of it that justify the study in the introduction but would leave the findings till later. (i.e. 'The analysis shows that the people...etc')

>\textcolor{blue}{moved comments about literature towards the section of concluding remarks; leaving some results in the introductory section}

About the structure of the paper: Make the paragraph L90-94 more interesting.

>\textcolor{blue}{Done.}

Context: The paragraph from 179 to 189  could be better grounded in evidence – do you have any relevant references about the transport situation in Santiago de Chile? – local council or local authority documents on transport, economic, planning situations etc.

>\textcolor{blue}{Recent social demands have highlighted the relevance and impacts of the neoliberal economic model in the development of the city – which correlates with the construction of the society and therefore shapes the development of transport infrastructure (see for recent protests and social crisis: Davies, 2019; Guardiola-Rivera, 2019; Sehnbruch, 2019). Even though everyday mobility in Santiago has also been contextualized in relation to other Latin American cities (Avellaneda and Lazo, 2011; Rodriguez Vignoli, 2008), there is much more to be explored in other regions in the Chilean context. Measuring instruments and new methodologies have allowed a more accurate and precise understanding of the social issues arising as consequence of transport infrastructure and housing provision (Cox and Hurtubia, 2016), and the minimum provision of basic services (Tiznado-Aitken et al., 2016). However, the focus on accessibility as a measure of inequities remains predominant (Tiznado-Aitken et al., 2018; Figueroa Martínez, et al., 2018; Niehaus et al., 2016; Rojas et al., 2016).}

Figure 1: Place figure 1 so that it is introduced in the text before the figure itself is presented. The figure could be formatted more attractively.

>\textcolor{blue}{Done.}

Survey description: L 200 The sections of the survey that were not used for the present study do not need to be listed.

>\textcolor{blue}{The study considers data from 3 out of 8 sections of a longer survey – data being more relevant for the study. The topics around these sections are Individual characteristics of respondents, feelings and emotions related to their commute, and aspects describing their regular commute trips.}

Results description: In general be more precise in the case study section – for example L210 – what percentage took commutes more than 60minutes.

>\textcolor{blue}{The sample tends to be younger (56% is younger than 34 years old), and well-educated (68% of respondents having technical or professional education), with an almost uniform distribution of income levels. The trend in typical commute time is towards longer commutes – e.g. 55% of the trips are longer than 40 minutes, and from those 56% corresponds to trips over the 60 minutes long.}

St Louis et al 2014: L157 Explain why ‘more research is needed’

>\textcolor{blue}{Because most literature focuses in trip satisfaction – the study goes beyond the understanding of determinants of trip satisfaction.}

Aggrupation of transport modes: P218 – why is motorcycle different to car? – they are both private motorised vehicles – explain the categories that you have split the modes into.

>\textcolor{blue}{Car (26% of total modes – which corresponds to 91% of the total of private motorised modes), Active (9% - Walking + Bicycle), Public (60% - Metro + Bus), and Other motorised modes (6% - Taxi + Colectivo + Motorcycle).}

Affective values: p.223 Why were these affective values chosen?  - needs to be explained. I would not normally think of ‘Poverty’ as a value, but rather as an economic condition, so explain the nature of these affective values.

>\textcolor{blue}{Words.}

Figure 2 – again make more attractive.

>\textcolor{blue}{Words.}

Who experiences dissonance: L257 Why is it important to know who experiences dissonance?

>\textcolor{blue}{Understanding who experiences dissonance is relevant in this study. We believe most disadvantageous groups will experience more dissonance with respect to their primary travel mode and affective values, namely freedom, enjoyment, happiness, feelings of poverty, luxury, and status.}

Figure 3: It may be me being a bit stupid but I found the Mosaic figures hard to understand – could you explain even more clearly? I can’t see tile J or K on figure 3. However, the ensuing discussion around which groups attribute which attributes to which mode is clear and well explained, and I grasped the general concept of what you were analysing, is just the figures I am unclear on.

>\textcolor{blue}{Words.}

\newpage

# Reviewer 2

Chile and wellbeing: This paper is about subjective well-being, mode of transport, and dissonance. In the introduction, three areas are identified where this study is expected to contribute. One contribution is that the study includes data from Chile, which is a country that has not been highlighted in previous transport research. This could be further discussed. Are there any thoughts/expectations of similar or different results from studies from other parts of the world? The authors partly address this by pointing to large income gaps, a different approach to transport but also a different culture. Perhaps this can be clarified.

>\textcolor{blue}{As noted in our response to Reviewer 1, recent social demands have highlighted the relevance and impacts of the neoliberal economic model in the development of the city – which correlates with the construction of the society and therefore shapes the development of transport infrastructure (see for recent protests and social crisis: Davies, 2019; Guardiola-Rivera, 2019; Sehnbruch, 2019). Even though everyday mobility in Santiago has also been contextualized in relation to other Latin American cities (Avellaneda and Lazo, 2011; Rodriguez Vignoli, 2008), there is much more to be explored in other regions in the Chilean context. Measuring instruments and new methodologies have allowed a more accurate and precise understanding of the social issues arising as consequence of transport infrastructure and housing provision (Cox and Hurtubia, 2016), and the minimum provision of basic services (Tiznado-Aitken et al., 2016). However, the focus on accessibility as a measure of inequities remains predominant (Tiznado-Aitken et al., 2018; Figueroa Martínez, et al., 2018; Niehaus et al., 2016; Rojas et al., 2016).}

Paragraph 3 is fuzzy and can be clarified. Remove redundant information and stick to the specific contribution. Then, compare points 1, 2 and 3 with the objectives presented later in the paper (e.g., line 247). It is important that the purpose is reproduced correctly throughout the paper.

>\textcolor{blue}{}

Relation of studies in the literature: The introduction provides a background to why subjective well-being is a relevant area to study. The authors highlight relevant references and show current knowledge within the field. However, different concepts are used and it is not always clear what is meant and how the different studies relate to each other. It is unclear whether previous studies focused on satisfaction, affective values, perceptions in general, or attitudes.

>\textcolor{blue}{}

Aims: This confusion also related to the aim of this specific paper, will it focus on affect or affective values (which may be something else?).

>\textcolor{blue}{}

Attitudes and satisfaction: Also in the introduction attitudes was mentioned as an important variable to study. It is also unclear why studies that focused on "life satisfaction" is discussed since this is not the focus of this study

>\textcolor{blue}{Because satisfaction has been the conventional way of addressing unconventional factors. Paragraph 3 clearly states the aim of the study focused on affective values – and also explain how this goes beyond satisfaction.}

Dissonance: I also think dissonance, which is an important concept in this study, is not discussed in the background.

>\textcolor{blue}{This has been discussed in section 3.2. paragraph 4.}

Taken together, the introduction needs clarification when it comes to previous studies, the aim of this specific study and the contribution to existing knowledge.

>\textcolor{blue}{See response to Reviewer 1.}

Although the authors highlight tools from transport research to capture well-being in travel, such as STS, it is not this instrument chosen for data collection. It is unclear why Freedom; Enjoyment; Happiness; Poverty; Luxury; and Status are selected concepts. Motivate the choice more clearly and how they relate to previous research.

>\textcolor{blue}{See response to Reviewer 1 about election of affective values.}

Developing and developed countries: In a concluding section, you choose to discuss the need for this type of research in developing countries. This could advantageously be a separate section where more clearly discusses and justifies any differences between developing and developed countries.

>\textcolor{blue}{This has been also commented in the Background section: “The research needs already recognized in the developed world are also markedly acute in the context of the Global South, where historical inequality has tended to create a symbolic attachment to the automobile, in addition to negative connotations for public transport and active travel”. A separated paragraph in the conclusion addresses this discussion, in the Chilean and LA context.}

Method: The data collection procedure is not described.

>\textcolor{blue}{The data collection considered a quota-sampling method for gathering the information, considering the socio-demographic information from Pre-Census of 2012. The survey was carried out face-to-face in centres of activity with dense provision of offices, services and educational centres.}

Separate results and analysis: I recommend separating results reporting and discussion/analysis to clarify what has been found.

>\textcolor{blue}{The results section has been divided in the two questions/aims of the study. The conclusion section discusses the findings.}

Figures 3, 4 and 7: The mosaic plot is however not helpful. Please consider other ways of presenting the data.

>\textcolor{blue}{Differs from Reviewer 3. The value of the mosaic figures is… we’ll keep them.}

Table 1: The tables are hard to interpret. I recommend to replace figure 1 with a table. Please be clear with the labels (for instance K-12 it unknown to me). If possible, add mean value and standard deviation to the table. This table should be in section 3.2.

>\textcolor{blue}{Words.}

The discussion is good although I miss a part on how this knowledge can help develop the alternatives and make them more attractive.

>\textcolor{blue}{See response to Reviewer 1.}

\newpage

# Reviewer 3

This paper is interesting as it details mode dissonance in terms of five affective values. The analysis method using visualization (through mosaic plots) is unique and suitable to the kind of data collected. Overall, the paper is structured very well.

>\textcolor{blue}{Words.}

SWB and health: The author(s) could at least add a paragraph about how SWB affects health, except for just mentioning that SWB has been found to be correlated with health.

>\textcolor{blue}{Words.}

Election of affective values: Second, the author(s) could make it clear how these affective values such as freedom, enjoyment, luxury and other are related to SWB. The connection is not quite clear in the first two sections.

>\textcolor{blue}{See our response to Reviewer 1 about election of affective values.}

Pg 6, L196-198: How were different areas divided?

>\textcolor{blue}{The data collection considered a quota-sampling method for gathering the information, considering the socio-demographic information from Pre-Census of 2012. The survey was carried out face-to-face in centres of activity with dense provision of offices, services and educational centres.}

Pg 217-218: Taxi, colectivo and motorcycle are grouped as same. I find that the affective evaluations of motorcycle users could be different than those made by users of the other two modes. It would be better if the author(s) could explain it or mention it in some part of their limitations.

>\textcolor{blue}{Addressed in Reviewer 1. Only 0,9% of the sample used motorcycle as primary transport mode.}

Figure 4: NA is shown in the legend but not in the figure. The x-axis labels are overlapped. I understand this might be due to low values.

>\textcolor{blue}{Words.}

Figure 7: Same as Figure 4. Is there any alternative way to make it better?

>\textcolor{blue}{Words.}

### 2020-08-14 Dissonance_Santiago_v2/Dissonance_Santiago_v2.Rmd
v2 BLM and AP revisions

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
   **Introduction**  
   Subjective wellbeing is a topic that has attracted considerable attention in recent years due to the way it correlates with health. From a transportation perspective, there is a burgeoning literature on the way travel can impact subjective wellbeing, and how this, in turn, can influence behavior.  
   **Objective**  
   The objective of this paper is to analyze a number of affective values associated with subjective wellbeing and the modes of transportation that people commonly use. In particular, we are interested in the potential for dissonance with respect to the primary mode of travel.  
   **Materials and Methods**  
   The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associate with the affective values of freedom, enjoyment, happiness, poverty, luxury and status. Analysis is based on tests of independence and visualization via mosaic plots.  
   **Results**  
   The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

   journal: "Journal of Transportation & Health"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article

@@

Transportation planning for decades has focused on providing mobility for the private automobile. This is a model of development that was initially introduced in North America as a solution to problems created by rapid urbanization, and that was eventually copied elsewhere [@Brown2009planning;@Angotti1996latin]. Despite the initial promise of automotive technology, it is now evident that mobility centred on the private automobile has given rise to a litany of maladies that are in urgent need of correction. This includes environmental concerns [i.e., climate change; @Chapman2007transport] as well as numerous other social [@Boschman2008toward; @Lucas2012transport; @Lucas2019evolution], health [@Khreis2016health; @Milne2012public], and equity issues [@Bocarejo2012transport; @Martens2012justice; @Pereira2017distributive].

As the real impacts of our societal dependence on the private autombobile have become increasingly evident, the transportation agenda has aimed to shift focus to the reduction of car use and towards the creation of mobility polycultures that offer a broader menu of transportation alternatives than primarily (or even just) the private automobile [@Miller2011collaborative]. In order to successfully achieve this goal, it is essential not only to provide the services and facilities that support public transportation and active travel, but also to attract new users to these modes of transportation [@Ettema2011satisfaction]. Within this context, it has been argued that sustainable transportation policies require all participants in the transportation system to challenge what Gossling and Cohen [-@Gossling2014sustainable] termed _transportation taboos_: deep-seated ideas concerning the contribution to emissions by individuals, the inequality of market-based approaches, and the social and psychological functions of transportation. With respect to the latter, it is important to move beyond a purely utilitarian focus if we are to understand how the affective value of the transportation experience can be leveraged to improve alternative transportation - particularly transit and active modes [@Anable2005work; @Domarchi2008effect; @Gatersleben2007affective; @Steg2005car].

A useful lens to investigate the social and psychological functions of transportation is by means of the concept of subjective wellbeing [SWB; see, inter alia, @Devos2013travel; and @Chatterjee2019commuting]. SWB is defined by the OECD as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” [@oecd2013oecd, p. 10]. As memorably put by Steg [-@Steg2005car]: is the use of car a must, or is it a lust? A close alignment, or consonance, between affective values and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode  (i.e., by satisfying the _lust_); on the other hand, dissonance, that is, the lack of concordance between the mode used and affective values would be detrimental to subjective wellbeing, especially of those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, one aim of the present research is to investigate _who_ experiences dissonance with respect to their primary mode of travel and affective values, namely freedom, enjoyment, happiness, feelings of poverty, luxury, and status. Furthermore, it is possible, even if use of a particular mode is a _must_, that travellers may still _lust_ for something else - in other words, the grass may in fact look greener from the window of the car. For this reason, we also aim to investigate which mode or modes are most commonly associated with affective values, with a focus on those travellers who experience dissonance in their primary mode.

Research is based on data collected from a sample of travellers in the city of Santiago in Chile. Survey respondents were asked about their primary mode of travel, and also about the mode or modes that they associate with the affective values mentioned above. The paper contributes to the literature in three ways, as follows:

1. The research reported here contributes to an emerging literature on the topic of transportation and affective values in the context of the Global South [@Alayyash2019commute; @Bejarano2017user; @Shao2019analysis; @Van2014effect; @Zorrilla2019exploring]; to the best of our knowledge, the case of Chile has not yet been reported.

2. Although there is an extensive literature on the enjoyment of commute and other affective values [see for instance @Paez2010enjoyment; @Redmond2001positive; @Whalen2013mode; @Ye2017satisfaction], from a hedonic and even eudaimonic perspectives the analysis has yet to be applied more fully in terms of distributional issues – i.e. which groups more commonly experience dissonance [see @Devos2018people].

3. The analysis shows the attitudes of people towards their primary mode and their perception towards ‘ideal modes’ – implying their preferences, even in situations when their ideal mode is not part of their actual choice set. More concretely, the results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status. We also find that there are some substantial variations in dissonance by age, education, income, and typical commute time. The attitudes of travellers towards transport modes are critical factors to be considered by policy-makers in case they want to promote and increase the use of public transport or active modes [@Bornioli2019affective; @Devos2017travel; @Devos2019satisfying; @Garling2019role; @Redman2013quality].

The structure of the paper is as follows. After these introductory remarks, we follow up with a background section that reviews the literature. Next, we discuss the case study and data used in the research. Then, the analysis and results are presented, before concluding with some discussion and directions for future research.

@@

A consensus has emerged in the transportation community regarding the need to complement the traditional utilitarian perspective of transportation by looking at mobility and transport issues from the lens of their affective functions. The affective value of transportation in turn is important due to its potential to improve or detract from SWB. One of the primary ways to explore this has been the satisfaction that travelers feel towards their every day mobility experience [e.g., @Bergstad2011subjective]. As a consequence, there is a wealth of research on satisfaction with the use of different modes of transportation. For example, numerous studies report that car users often have a higher level of satisfaction compared to other transport modes [@Bergstad2011affective; @Eriksson2013perceived; @Redmond2001positive; @Whalen2013mode; but see @Handy2019commute]. In a similar way, there are multiple reports that active travel also tends to yield high levels of satisfaction [@Gatersleben2007affective; @Smith2017commute; @Stlouis2014happy; @Paez2010enjoyment; @Whalen2013mode; @Handy2019commute], In contrast, public transport users tend to assess their experience more negatively [@Abenoza2017travel; @Devos2016travel; @Gatersleben2007affective; @Handy2019commute; @Paez2010enjoyment]. Multi-modal trips also influence satisfaction levels; for instance, when an individual chooses a particular mode of transportation, she will report a higher level of satisfaction with that chosen mode – perhaps as a form of _post hoc_ validation [@Susilo2014exploring].

While the use of travel satisfaction has been mainly used in the context of daily trips – typically linked to cost-benefit and utilitarian measurements –, the evaluation of Subjective Wellbeing (SWB) over time has risen as an alternative measure. In the field of travel behaviour, Ettema et al. [-@Ettema2010out, p. 725] define SWB as the degree to which an individual positively evaluates the overall quality of their lives, where the general life satisfaction encompasses a more extended temporality – which implies assuming a tendency to be more stable over time. This concept has prompted a growing literature that complements and applies SWB in a broader range of satisfaction scales and situations. The definition of other factors such as travel choice mode, attitudes and external elements of the built environment have been studied for a broader understanding of the changes produced in the SWB [e.g., @Handy2019commute]. As these factors do not necessarily apply to the general life satisfaction on the long term, the studies have aimed to determine both the direct and indirect effects on the perception of users [see, e.g., @Ye2017satisfaction]. Other concepts have also emerged as the Satisfaction with Travel Scale (STS), a measurement devised by Ettema et al. [-@Ettema2011satisfaction], as well as different scales based on people’s travel perceptions. De Vos et al. [-@Devos2015satisfying], for instance, explore in detail the underlying dimensions of the affective domain of STS on which SWB is based [for more on STS see also @Friman2013psychometric].

Recent literature on SWB and its link with transport have demonstrated a relationship between people’s perceptions and satisfaction with their daily travel [e.g., @Smith2017commute; @Stlouis2014happy; @Mokhtarian2018travel]. Scholars have shown that accessibility has been the most developed factor that influences people’s wellbeing [@Delbosc2012role], and activities have a direct impact on travel satisfaction [@Bergstad2011subjective]. Delbosc [-@Delbosc2012role, p. 28], for instance, has summarised the most significant influences on psychological wellbeing: poverty and employment, meaningful relationships and health. However, understanding the components affecting people’s perceptions implies the differentiation between affective (also named as symbolic-affective) and instrumental values [@Bergstad2011affective]. Steg et al. (2011) have compared symbolic-affective opposed to instrumental-reasoned motives based on car-use, and other studies have also found associations between affective and symbolic aspects of car-use [see, e.g., @Gatersleben2007affective; @Lois2009relationship]. Previous studies have also demonstrated how socio-demographic factors affect the levels of SWB. The effect of income on SWB [@Clark1996satisfaction; @Ferrer2005income]; education and unemployment [@Argyle1999causes]; age [@Diener1997subjective], and gender [@Tesch2008gender] have been extensively studied. Recent research also suggests the links between commuting, SWB and emotional wellbeing assessment (**Olson et al. 2013**; **Kahneman et al. 2004**). However, more research is needed to understand how these socio-demographic variables connect as well with the affective responses to mode of travel [@Stlouis2014happy].

The research needs already recognized in the developed world are also markedly acute in the context of the Global South, where historical inequality has tended to create a symbolic attachment to the automobile, in addition to negative connotations for public transport and active travel [@Zorrilla2019exploring]. In this way, there is an emerging literature that investigates affective factors in travel behaviour in a number of developing countries. A cross-country study in Asia revealed that the affective factors of public transportation and car use are important, and in particular the social orderliness of transit was suggested as a way to make this mode more attractive to users [@Van2014effect]. In terms of active travel, a study in China found that attitudes that embrace new styles and technologies despite their cost are associated with the intention to continue using shared bicycles [@Shao2019analysis]. The importance of affective factors for policy and planning is further highlighted by research in Colombia that shows how users felt proud using a bicycle shared system, in addition to experiencing feelings of belonging to a civic culture and the enjoyment and pleasure of travel itself [@Bejarano2017user]. This paper contributes to further our understanding of affective values in travel behaviour in a developing country.

@@

The case study to contextualize the discussion is Santiago de Chile, the capital of one of the countries with the highest levels of inequality in terms of socio-economic distribution in the world. These inequities have manifested in a disproportionate expense in transport by the different socioe-conomic sectors, a disproportion in travel times and distances travelled. On the other hand, car ownership and use of public transport in the capital are directly conditioned by strong spatial segregation – the higher the income, the higher is the use of the automobile; conversely, and the lower the income, the higher is the predisposition to use public transport. Although transport is a reflection of deep inequities in various dimensions, it has triggered an exceptional discomfort and disagreement with the daily travel experience and quality of life.

## Sample

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
   *Introduction*  
   Subjective wellbeing is a topic that has attracted considerable attention in the transportation literature in recent years. As a result, there is a burgeoning literature that investigates the impacts of travel on subjective wellbeing, and how wellbeing, in turn, can influence behavior. An important aspect of subjective wellbeing are the affective reactions of people to their experiences.   
   *Objective*  
   The objective of this paper is to analyze the affective reactions of travelers with respect to various modes of transportation. In particular, we are interested in the potential for dissonance between primary mode of travel and the mode(s) of travel identified as evoking various affective reactions.  
   *Materials and Methods*  
   The study is based on data collected from a sample of travelers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name their ideal mode(s) of transportation from the perspective of various affective reactions. The reactions we investigate are associated with the values of freedom, enjoyment, happiness, poverty, luxury, and status. Analysis is based on tests of independence and visualization techniques.  
   *Results*  
   The results indicate that users of public transportation experience the most dissonance in terms of affective reactions, and active travelers the least. For those travelers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, whereas public transportation is most commonly associated with poverty. The automobile, in contrast, is the mode most commonly associated with luxury and status.

journal: "Journal of Transport & Health"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article

@@

Transportation planning for decades has focused on providing mobility for the private automobile. This is a model of development that was initially introduced in North America as a solution to problems created by rapid urbanization, and that was eventually copied elsewhere [@Brown2009planning;@Angotti1996latin]. Despite the initial promise of automotive technology, it is now evident that mobility centered on the private automobile has given rise to a litany of maladies that are in urgent need of correction. This includes environmental concerns [i.e., climate change; @Chapman2007transport] as well as numerous other social [@Boschman2008toward; @Lucas2012transport; @Lucas2019evolution], health [@Khreis2016health; @Milne2012public], and equity issues [@Bocarejo2012transport; @Martens2012justice; @Pereira2017distributive].

As the impacts of our societal dependence on the private automobile have become increasingly evident, the transportation agenda has aimed to shift focus to the reduction of car use and towards the creation of mobility polycultures that offer a broader menu of transportation alternatives other than primarily (or even just) the private automobile [@Miller2011collaborative; @Lavery2013driving]. In order to successfully achieve this goal, it is essential not only to provide the services and facilities that support public transportation and active travel, but also to attract new users to these modes of transportation [@Ettema2011satisfaction]. Within this context, it has been argued that sustainable transportation policies require all participants in the transportation system to challenge what Gossling and Cohen [-@Gossling2014sustainable] termed _transportation taboos_: deep-seated ideas concerning the contribution to emissions by individuals, the inequality of market-based approaches, and the social and psychological functions of transportation. With respect to the latter, subjective wellbeing (SWB) is increasingly recognized as a policy goal in and of itself [@Diener2015national; @Dolan2012measuring], and also as an intermediate objective to achieve ulterior goals. Accordingly, there is a growing consensus in the field of transportation about the need to move beyond a purely utilitarian focus to understand the affective and wellbeing values of the transportation experience [e.g., @Anable2005work; @Devos2013travel; @Domarchi2008effect; @Gatersleben2007affective; @Steg2005car]. Consideration of factors related to subjective wellbeing, for example, are useful for policy-makers to enhance the experience of travel, to increase satisfaction with travel, and to meet the preferences of travelers [@Chatterjee2019commuting].

A definition of subjective wellbeing is as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” [@oecd2013oecd, p. 10]. As memorably put by Steg [-@Steg2005car]: is the use of a car a must, or is it a lust? A close alignment, or consonance, between affective reactions and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode (i.e., by satisfying the _lust_); on the other hand, dissonance, namely the lack of concordance between the mode used and the affective reaction to that mode, would be detrimental to subjective wellbeing, especially to those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, the present research aims to investigate _who_ experiences affective dissonance with respect to their primary mode of travel. We focus our investigation on reactions from the perspective of feelings of freedom, enjoyment, happiness, poverty, luxury, and status - concepts termed _affective values_ hereafter. Furthermore, it is possible, even if use of a particular mode is a _must_, that travelers may still _lust_ for something else - in other words, the grass may look greener from the window of a car. For this reason, we also aim to investigate which mode or modes are most commonly identified as ideal from the perspective of the aforementioned affective values.

Our research is based on data collected from a sample of travelers in the city of Santiago in Chile. The analysis sheds light on the affective reactions of people towards their primary mode of travel, and their perceptions towards 'ideal modes'. This reveals important differences in situations when the ideal mode of a traveler is not the mode they actually use - or in other words, the affective dissonance between actual and ideal modes of travel. More concretely, the results indicate that users of public transportation experience the most dissonance, followed by automobile users, and finally active travelers, who tend to experience the least affective dissonance. For those travelers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, whereas public transportation is most commonly associated with poverty, and the automobile is most commonly associated with luxury and status. We also find that there are significant variations in dissonance by age, education, income, and typical commute time. The reactions of travelers towards various transport modes are critical factors that policy-makers need to consider to promote and increase the use of public transport and active modes [@Bornioli2019affective; @Devos2017travel; @Devos2019satisfying; @Garling2019role; @Redman2013quality].

After these introductory remarks, the paper follows up with a review of the literature on the topics of subjective wellbeing and affective values. Next, we discuss the case study, as well as the methodology and data used for this research. The analysis and results of the study are presented afterwards, before concluding with some discussion and directions for future research and policy.

Please note that this document  was prepared using R Markdown and is an example of reproducible research [@Brunsdon2020opening]. The R Markdown file, along with the data file needed to reproduce the analysis, are available for download\footnote{\url{https://drive.google.com/open?id=189ZvfVvRis5xZA9IlviKCLW2bwbPJAgg}}.

@@

A consensus has emerged in recent years regarding the need to complement the traditional utilitarian perspective of transportation by investigating mobility and transport from the lens of their affective functions. The affective value of transportation is important due to its potential to improve or detract from SWB.

One of the primary ways to explore issues of subjective wellbeing in transportation has been the satisfaction that travelers feel towards their everyday mobility experience [e.g., @Bergstad2011subjective]. As a consequence, there is a wealth of research on satisfaction with the use of different modes of transportation. For example, numerous studies report that car users often have a higher level of satisfaction compared to other transport modes [@Bergstad2011affective; @Eriksson2013perceived; @Redmond2001positive], although others report highest commute satisfaction for bicycle and train commuters [e.g., @Whalen2013mode; @Handy2019commute]. In a similar way, there are multiple reports that active travel also tends to yield higher levels of satisfaction [@Gatersleben2007affective; @Smith2017commute; @Stlouis2014happy; @Paez2010enjoyment; @Whalen2013mode; @Handy2019commute], In contrast, public transport users often assess their experience more negatively [@Abenoza2017travel; @Devos2016travel; @Gatersleben2007affective; @Handy2019commute; @Paez2010enjoyment]. Multi-modal trips also influence satisfaction levels; for instance, individuals tend to report higher levels of satisfaction with the mode they use – perhaps as a form of _post hoc_ validation of their choices or circumstance [@Susilo2014exploring].

While travel satisfaction has often been used in the context of daily trips – typically linked to cost-benefit and utilitarian measurements –, the evaluation of subjective wellbeing (SWB) over time has risen as an alternative measure. Ettema et al. [-@Ettema2010out, p. 725] define SWB as the degree to which an individual positively evaluates the overall quality of their lives, where the general life satisfaction encompasses a more extended temporal period – which implies greater temporal stability. This concept has prompted a growing literature that complements and applies SWB in a broader range of satisfaction scales and situations. The definition of other factors such as travel choice mode, attitudes, and external elements of the built environment have been studied for a broader understanding of the changes produced in the SWB [e.g., @Handy2019commute]. As these factors do not necessarily apply to longer-term general life satisfaction, other studies have aimed to determine both the direct and indirect effects on the perception of users [see, e.g., @Ye2017satisfaction]. Other concepts have also emerged, including the Satisfaction with Travel Scale [@Ettema2011satisfaction], as well as different scales based on people’s perceptions of travel. De Vos et al. [-@Devos2015satisfying], for instance, explore in detail the underlying dimensions of the affective domain of STS on which SWB is based [for more on STS also see @Friman2013psychometric].

The literature on SWB has demonstrated a relationship between people’s perceptions and satisfaction with their daily travel [e.g., @Smith2017commute; @Stlouis2014happy; @Mokhtarian2018travel]. Scholars have shown that accessibility is among the most studied factors that influences subjective wellbeing [@Delbosc2012role], and activities have a direct impact on travel satisfaction [@Bergstad2011subjective]. Delbosc [-@Delbosc2012role, p. 28], for instance, has summarized the most significant influences on psychological wellbeing: poverty and employment, meaningful relationships and health. However, understanding the components that affect these perceptions implies the differentiation between affective (also named as symbolic-affective) and instrumental values [@Bergstad2011affective]. Steg et al. (2011) have compared symbolic-affective and instrumental-reasoned motives for car-use, and similar findings are reported in other studies [e.g., @Gatersleben2007affective; @Lois2009relationship]. Previous research also demonstrates how socio-demographic factors affect the levels of SWB, includin income [@Clark1996satisfaction; @Ferrer2005income], education and unemployment [@Argyle1999causes], age [@Diener1997subjective], and gender [@Tesch2008gender]. However, further research beyond the determinants of trip satisfaction is needed to understand how these socio-demographic variables connect with the affective reactions to various modes of travel [@Stlouis2014happy].

The research needs outlined above are well-recognized in the developed world, but there is still a dearth of research in the context of the Global South. Historical inequality in many developing countries has led to strong symbolic attachments to the automobile, in addition to negative connotations for public transport and active travel [@Zorrilla2019exploring]. To address this gap in knowledge a body of literature has emerged to investigate the affective aspects of travel behavior in a number of developing countries. A cross-country study in Asia revealed that the affective factors of public transportation and car use are important, and in particular the social orderliness of transit was suggested as a way to make this mode more attractive to users [@Van2014effect]. In terms of active travel, a study in China found that attitudes that embrace new styles and technologies despite their cost are associated with the intention to continue using shared bicycles [@Shao2019analysis]. The importance of affective factors for policy and planning is further highlighted by research in Colombia that shows the pride users feel when using a bicycle shared system. This affective reaction is above and beyond other positive feelings, such as a sense of belongingness to a civic culture, and the enjoyment and pleasure of active travel itself [@Bejarano2017user].

@@

The case study for the present research is Santiago. Santiago is the capital of Chile, a country with one of the highest levels of socio-economic inequality in the world. A tangible manifestation of the inequalities experienced by many in Chile is the large disparity in the relative cost of transportation, travel time, and distance traveled by different socio-economic segments of the population. To further complicate matters, strong spatial segregation also conditions car ownership and the use of public transportation – the higher the income, the higher the use of the automobile; conversely, the lower the income, the higher the reliance on public transportation. Although the transportation experience is but one of many dimensions of inequality, the experiences in this sector have triggered exceptional discomfort and dissatisfaction among the public. Recent social unrest, triggered by a seemingly minor hike in the fare of public transportation, has brought many of these concerns to the forefront of the public conscience in dramatic fashion [e.g., @Davis2019; @GuardiolaRivera2019; @Sehnbruch2019].
